**Microbiota predict _Clostridium difficile_ severity in germ-free mice colonized with human feces**

### 2016-11-15 submission/Manuscript.Rmd

**Susceptibility to infection is community-dependent.** _C. difficile_ strains are known to vary in disease phenotypes based both on pathogenicity factors of the strain and host microbiome composition (cite someone). To test the susceptibility of human-derived microbiomes to _C. difficile_ infection, two donor-mouse microbiomes that were previously shown to be resistant to _C. difficile_ strain 431 induced weight loss (Fig 2B) (DA00396, DA00430) and one that was suscpetible to severe disease (DA00578) were independently challenged with one of three unique _C. difficile_ strains (table or reference for toxin prod, etc). As done previously, _C. difficile_ stool CFU was enumerated daily and mouse weight-loss tracked. Infection of mice with resistant communities with virulent strain 458 or the weaker strain 299 resulted in colonization to levels near that of 431 (Fig 5A) and these mice did not experience severe weight loss or show clinical signs of disease (Fig 5B). In contrast, infection of mice harboring microbiomes susceptible to severe disease (DA00578, Fig 1B) with strain 3

**Results to be written**

2.	Propensity for severe CDI is community-dependent and strain-independent
a.	Figure 5:  Infection of mice with different C. difficile strains.

### Discussion
•	Restate results,
•	caveats about mouse weights 
•	No donors were colonization resistant, discuss donor differences 
severity, asymptomatic colonzation discussion? 
•	Discuss prediction methods and outcomes
•	Discuss potential mechanisms for interesting OTUs
•	Discuss different strain results
•	Future work blah blah

@@

**Human samples** Donor stool samples were chosen from a sample bank collected from patients from October 2010 to November 2012 at the University of Michigan. Sample collection methods and study was approved by the Institutional Review Board at the University of Michigan. Stool samples from University of Michigan Health System patients were submitted to the microbiology lab for _C. difficile_ testing. _C. difficile_ presence was measured using WHATEVER- Cdiff quik check? toxin AB?. Samples from these patients were stored in Cary-Blair medium at -80C. All patients consented to participation in the study. Donor stool for this study was selected from one patient that tested positive for _C. difficile_ infection and four patients that presented with diarrhea that tested negative for _C. difficile_. Healthy donor stool was collected from volunteer donors living in the Ann Arbor, MI area.

>>>>

**Susceptibility to infection is community-dependent.** _C. difficile_ strains are known to vary in disease phenotypes based both on pathogenicity factors of the strain and host microbiome composition (cite someone). To test the susceptibility of human-derived microbiomes to _C. difficile_ infection, two donor-mouse microbiomes that were previously shown to be resistant to _C. difficile_ strain 431 induced weight loss (Fig 2B) (DA00396, DA00430) and one that was suscpetible to severe disease (DA00578) were independently challenged with one of three unique _C. difficile_ strains (table or reference for toxin prod, etc). As done previously, _C. difficile_ stool CFU was enumerated daily and mouse weight-loss tracked. Infection of mice with resistant communities with virulent strain 458 or the weaker strain 299 resulted in colonization to levels near that of 431 (Fig 5A) and these mice did not experience severe weight loss or show clinical signs of disease (Fig 5B). In contrast, infection of mice harboring microbiomes susceptible to severe disease (DA00578, Fig 1B) with strain 395 resulted in colonization and severe weight loss by two days post-infection (Fig 5A, 5B). Strain 395 is similarly virulent as strain 431, and shows a similar infection phenotype in these mice. These preliminary results suggest that human-associated microbiomes may be more predictive of disease severity than the _C. difficile_ strain.

### Discussion

Study of human-associated microbiomes and their interactions with _C. difficile_ has been limited by the lack of an appropriate model and available samples. Here we colonized 16 groups of mice with human donor stool from patients with different clinical profiles or healthy volunteers. This created a diverse set of human-associated microbiomes in the mice that allowed for testing _C. difficile_ infection dynamics and community interactions. Our results showed that both healthy and diarrheal-associated human communities were susceptible to severe disease. Additionally, our results confirmed our and others previous studies that proved that resistance to _C. difficile_ disease is dependent on the entire microbial community and not just one or two species (cite alyx). Further, we were able to learn classification models that identified specific OTUs within susceptible communities that predispose mice to infection. By subsequently infecting susceptible or resistant communities with varying strains of _C. difficile_ we confirmed that the initial state of the human-associated microbiome is more predictive of disease severity than the strain used.

Here we were able to colonize germ-free mice with a variety of human stool communities from diarrheal or healthy people. After allowing these communities to colonize the mice for 2 weeks, they were challenged with _C. difficile_. It is important to note that the mouse gut environment is different from that of the human, and as such exerts a selective pressure on the human stool microbial community that stabilizes over the 2 weeks. Thus these mice cannot be considered "humanized" but rather containing human-associated microbes. Interestingly, clinical status was not associated with proclivity to severe _C. difficile_ infection as some healthy donor mice were susceptible. A surprising finding was that none of the human-associated mouse communities were completely resistant to colonization by _C. difficile_. This is in contrast to numerous reports of mouse microbiome studies (cite) and the knowledge that in general healthy humans without gut perturbations are resistant to _C. difficile_ infection (cite). On the other hand, it is known that humans, especially health care employees, can be asymptomatic carriers of _C. difficile_ (though at what CFU?!). This may explain the fact that the majority of our groups of mice were colonized by _C. difficile_ but did not display outward signs of illness or infection-associated weight loss (Fig 2).

•	No donors were colonization resistant, discuss donor differences 
severity, asymptomatic colonzation discussion? 
•	Discuss prediction methods and outcomes - Nick
•	Discuss potential mechanisms for interesting OTUs - Nick
•	Discuss different strain results
•	Future work blah blah

@@

**Human samples** Donor stool samples were chosen from a sample bank collected from patients from October 2010 to November 2012 at the University of Michigan. Sample collection methods and study was approved by the Institutional Review Board at the University of Michigan. Stool samples from University of Michigan Health System patients were submitted to the microbiology lab for _C. difficile_ testing. _C. difficile_ presence was measured using WHATEVER- Cdiff quik check? toxin AB?. Samples from these patients were stored in Cary-Blair medium at -80C. All patients consented to participation in the study. Donor stool for this study was selected from one patient that tested positive for _C. difficile_ infection and five patients that presented with diarrhea and tested negative for _C. difficile_. Healthy donor stool was collected from adult volunteers living in the Ann Arbor, MI area. 100ul of stool slurry was gavaged into each mouse.

**_Clostridium difficile_ and sample processing** For all experiments, 100 spores of _Clostridium difficile_ were used to orally infect mice as described previously (cite alyx). The _C. difficile_ strains used were acquired from spore stocks of (WHERE) and comprised of strains 431, 395, 299 and 468 (cite?). After gavage, the remaining inoculum was diluted and plated to confirm infection dose. Bacteria were cultured anaerobically on TCCFA plates and incubated at 37C. 
During infection, two fresh stool pellets were collected from each mouse daily. One pellet was resuspended in anaerobic phosphate-buffered saline and diluted and plated on TCCFA plates to enumerate _C. difficile_ CFU. The other pellet was frozen at -80C until the end of the experiment.

**DNA isolation and sequencing** Bacterial DNA was isolated from banked donor stool and each mouse stool pellet using the MO-BIO PowerSoil DNA isolation kit. This DNA was used as template for
